---
title: "Joe_Rovalino_Data607_wk#5"
author: "Joe Rovalino"
date: "9/29/2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Relevant Information:
The chart was loaded into MySQL DB and describes arrival delays for two airlines across five destinations. Your task is to:
(1) Create a .CSV file (or optionally, a MySQL database!) that includes all of the information above. - CHOSE to load to MySQL and use the lesson from homework 2 to create DB.
You’re encouraged to use a “wide” structure similar to how the information appears above, so
that you can practice tidying and transformations as described below.
(2) Read the information from your .CSV file into R, and use tidyr and dplyr as needed to tidy
and transform your data.
(3) Perform analysis to compare the arrival delays for the two airlines.
(4) Your code should be in an R Markdown file, posted to rpubs.com, and should include narrative
descriptions of your data cleanup work, analysis, and conclusions. Please include in your
homework submission:
The URL to the .Rmd file in your GitHub repository. and
The URL for your rpubs.com web page.

```{r}
library(getPass)
library(RMySQL)

db_user <- 'root'
db_password <- getPass::getPass("Enter the password: ")
db_name <- 'data607wk5'
db_table <- 'fltbycity'
db_host <- '127.0.0.1' # for local access
db_port <- 3306

mydb <-  dbConnect(MySQL(), user = db_user, password = db_password,
                 dbname = db_name, host = db_host, port = db_port)

s <- paste0("select * from ", db_table)
rs <- dbSendQuery(mydb, s)
df <-  fetch(rs, n = -1)
on.exit(dbDisconnect(mydb))

df
```
#write to CSV file for upload to grading site. Will also upload sql script used to create the DB #table.
# good site http://sphweb.bumc.bu.edu/otlt/MPH-Modules/BS/R/R-Manual/R-Manual5.html

```{r}
write.csv(df, 'fltbycity.csv',row.names=FALSE)
```
# https://tibble.tidyverse.org/
# Tidy work. Use Control Shift M for #shortcut to pipes

```{r}
library(tidyverse)
library(dplyr)

fltdata <- as_tibble(df)
fltdata

#check gahter worked on  
#fltdatachk <- fltdata %>% gather(city, count, -id, -airline, -time_perf)
# select gets rid of id field
#spread to widen the time performance column

fltdata2 <- fltdata %>% gather(city, count, -id, -airline, -time_perf) %>% select (airline, time_perf, city, count) %>%  spread (time_perf, count)

fltdata2

# Add Percent on time to data frame fltdata2 and total count

fltdata3 <- fltdata2 %>% mutate( percontime = on_time/(on_time + delayed), total_flights = (on_time + delayed))
fltdata3

  ggplot(fltdata3, aes(fill=airline, y=percontime, x=city)) +
    ggtitle("Ontime by City") +
    theme(plot.title = element_text(hjust = 0.5)) + 
    geom_bar(position='dodge', stat="identity") +
    xlab('City') + 
    ylab('Percent of Flights On time')
```

Conclusion:
1) Alaska airlines has less delays in each airport
2) Philly appears to be the most on time of all the airports for both airlines from a percentage perspective
3) San Francisco seems to be the most delayed airport from the graph for both airlines. 
4) I would fly Alaska if I were concerned with being at an airport on time. 